---
title: "Asking for help"
author: "Emily O'Grady"
date: "January 22 2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
This is a session designed to help you ask for help, in a consistent and efficient way that
will facilitate better discussion and faster solutions. It is designed for Science
staff, but could be relevant to a broader audience. 

Everyone experiences issues and bugs when navigating and developing code. The most 
important tool you can have is the ability to ask for help. This session will cover
the best practices for asking for help, and how to engage with AI to get the best
results.

# Platforms

Where to go when you run into a problem? Here are some of the most popular platforms
for asking for help:

- *Stack Overflow* (https://stackoverflow.com/)
  Usually the first thing to come up in your google searches. It is a forum for devlopers to ask
  and answer questions. It is a great resource for finding solutions to common problems.
  Use with some caution, as solutions may not always be the best practice and there may be nuances 
  to your problem, due to the security of your DFO laptop.
- *GitHub* (https:://github.com)
  A great place to find code and packages which may help with your work. On 
  individual packages, the 'issues' boards can be a great place to ask for
  help with a niche problem with a particular library.
- *AI* (https:://m365.cloud.microsoft/chat)
  There are a number of AI tools which can help you debug your code. These are
  great for getting a quick solution to a problem, but may not always be the best
  practice. The recommended tool for DFO users is Bing AI (Copilot). Github copilot
  is also now available and can be used in most IDEs. 
- *DFO channels*
  The Science PL community channel on Teams is active and a great place to go with 
  questions that may have specific DFO nuances, issues with our internal databases
  or other problems that involve coordination with IT. 


# Making a reproducible example

A reproducible example is key to helping someone understand your problem. 
This will allow someone else to run your code, and troubleshoot it in their own environment. 
It makes it much easier to get a clear answer to your problem.

Here are some references:
https://stackoverflow.com/questions/5963269/how-to-make-a-great-r-reproducible-example
https://stackoverflow.com/help/minimal-reproducible-example

Let's walk through some functions that can be helpful in generating a reproducible example.

Particularly if you cannot share your data. 

```{r}
# generate a random dataframe
# using rnorm or set.seed 
df <- data.frame(
  x = rnorm(100),
  y = rnorm(100)
)

# grab an example dataset (built into R)
data(mtcars)

# grab an example oceanographic dataset
library(ocedata)
# ?ocedata
# CHALLENGE: find a dataset that could represent data used in your work
# data()

```

The next step in making a reproducible example is to isolate and simplify your code.
This will help you identify the problem, and make it easier for someone else to help you.
You should be including the code as well as your expected output, and the actual output.

You can also consider formatting the code within fences to make it easier to read.
Make sure that you are including all the libraries you have used, and that the code is
clearly commented.

```{r eval = FALSE}
# Let's try to tidy this up before we debug it

data(secchi)
data <- secchi

data_2 <- data %>%
  filter(depth < 10)

data_3 <- data_2 %>%
  group_by(latitude, longitude) %>%
  summarise(mean_depth = mean(depth))

mdi <- grep(colnames(data_3), 'mean_depth')

data_3[mdi] <- round(data_3[mdi], 2)

data_3$station <- 'HL_02'
data_3$station[1000:4000] <- 'HL_03'
data_3$station[4000:5939] <- 'HL_04'

ggplot(data_3, aes(x = station, y = mean_depth)) +
  geom_bar()

```


# Getting details

Sometimes people will need more details to help you! 
Sometimes these details can help you as well.
Here are some things you can use:

- *sessionInfo()*
  This will show which version of R you are using, and which packages are loaded.
  
- *browser()*
  This will pause your code at a certain point, and allow you to inspect the variables at that point.
  
  For example:
  
```{r}
  df <- data.frame(
  x = rnorm(100),
  y = rnorm(100)
)

  for (i in seq_along(df$x)) {
    if (df$x[i] < 0) {
      browser()
    }
    print(df$x[i])
  }
```
  
- *print statements and logs*
  You can use print statements to see what is happening at a certain point in your code.
  You can also write logs to a file, to see what is happening over time.
  
  For example:
  
```{r}
 df <- data.frame(
  x = rnorm(100),
  y = rnorm(100)
)
sink('error_log.txt')
  for (i in seq_along(df$x)) {
    if (df$x[i] < 0) {
      print(paste('Value is less than zero! (i = ', i, ')'))
      # browser()
    } else{
    print(df$x[i])
    }
  }
sink()

```

# Tips for engaging with AI
- Be specific.
- Use code snippets. This will help the AI understand your problem, and is as easy as copy-paste!
- Narrow your libraries, if you would like the answer to only include specific libraries
- Copy the error message from your console
- Give it solution ideas or direction. Sometimes prompting can help
  jumpstart solutions
- Explain your data, outline columns names and data types
- Ask solutions to follow 'best practices'


https://octawebtools.com/mastering-chatgpt-for-coding-tips-tricks-and-detailed-examples/
https://medium.com/@pranshu1902/code-debugging-made-easy-with-chatgpt-your-personal-debugging-assistant-c9f1adc054c8
https://rollbar.com/blog/how-to-debug-code-using-chatgpt/
https://www.cyfrin.io/blog/ai-code-debugger-debug-code-with-ai



# Cheat sheets

Cheat Sheets can be a great quick reference to help you with your coding
on a day to day basis. Here are some of the most popular cheat sheets for R:

cmocean colormaps: https://www.researchgate.net/figure/Colormaps-available-in-the-cmocean-package_fig2_307517997
posit: https://rstudio.github.io/cheatsheets/
base graphics: https://publish.illinois.edu/johnrgallagher/files/2015/10/BaseGraphicsCheatsheet.pdf
base R: https://geomoer.github.io/moer-base-r/cheatsheet.html
r for ecology: https://www.rforecology.com/post/the-essential-functions-of-r-cheatsheet/

# Example

This code is not working. Can you help me debug it?

The goal is to find locations which were sampled more than twice and 
then generate a mean of all measurements for that station. We then want to plot
only stations where the mean depth is greater than 5.

```{r, eval = FALSE}
library(tidyverse)
data(secchi)

# find locations which were sampled more than twice
stations <- secchi %>%
  group_by(c(latitude, longitude)) %>%
  filter(n > 2) %>%
  summarise(mean = mean(depth))

# filter lat and lon data where the mean is greater than 5
stations <- stations %>%
  filter(mean > 5)

# name the stations with unique IDs
stations$station <- seq_along(stations$latitude)

# plot
ggplot(stations, aes(x = station, y = mean)) +
  geom_col()


```
