---
title: "Downloading Oceanographic Data from ERDDAP"
author: "Your Name"
date: "2024-12-18"
output: html_document
---

## Introduction

In this example, we will use the `rerddap` package to download and analyze oceanographic data from an ERDDAP server. ERDDAP is a data server that provides a simple, consistent way to download subsets of scientific datasets.

## Setup

First, we need to install and load the necessary packages:

```{r setup, include=FALSE}
# Install rerddap if not already installed
if (!requireNamespace("rerddap", quietly = TRUE)) {
  install.packages("rerddap")
}
library(tidyverse)
library(rerddap)
library(ggplot2)
library(cmocean)
```

## Downloading data

We will download sea surface temperature (SST) data from the NOAA ERDDAP server. The dataset ID for this example is erdMBsstdmday.

```{r download-data}
# Define the dataset ID and the time range
dataset_id <- "erdMBsstdmday"
time_range <- c("2023-01-01", "2023-06-31")

# Use the griddap function to download the data
sst_data <- griddap(
  datasetx = dataset_id,
  time = time_range,
  latitude = c(30, 40),
  longitude = c(130, 120),
  fields = "sst"
)

# Display the structure of the downloaded data
str(sst_data$data)
```

## Visualizing Data

Next, we will visualize the downloaded SST data using ggplot2.

```{r visualize-data}
# Convert the data to a data frame
sst_df <- as.data.frame(sst_data$data)

# average by month
sst_df <- sst_df %>%
  dplyr::mutate(month = format(as.POSIXct(sst_df$time), '%m')) %>%
  dplyr::group_by(month) %>%
  dplyr::summarise(monthly_avg_sst = mean(sst, na.rm = TRUE))

# Plot the SST data
ggplot(sst_df, aes(x = month, y = monthly_avg_sst)) +
  geom_col( ) +
  labs(title = "Monthly Average Sea Surface Temperature (SST)",
       x = "Month",
       y = "SST [deg C]") +
  theme_minimal()
```
